---
title: "Week 4"
author: "Emily Esposito"
date: "10/21/2021"
output: html_document
---
# Codebook
We have a lot of columns, here's what they are:
PARTY: Political Party
demFav: Average favorability towards Democrats
repFav: Average favorability towards Republicans
Q2A_A: Donald Trump, R
Q2A_B: Bill Weld, R
Q2A_C: Nancy Pelosi, D
Q2A_D: Alexandria Ocasio-Cortez, D
Q2A_E: Mitch McConnell, R
Q2A_F: Barack Obama, D
Q2B_A: Joe Biden, D
Q2B_B: Cory Booker, D
Q2B_C: Pete Buttigieg, D
Q2B_D: Kamala Harris, D
Q2B_E: Bernie Sanders, D
Q2B_F: Elizabeth Warren, D
Q2B_G: Steve Bullock, D
Q2B_H: Julian Castro, D
Q2B_I: Bill de Blasio, D
Q2B_J: Tulsi Gabbard, D
Q2B_K: Kirsten Gillibrand, D
Q2B_L: Amy Klobuchar, D
Q2B_M: Beto O'Rourke, D
Q2B_N: Tom Steyer, D
Q2B_O: Marianne Williamson, D
Q2B_P: Andrew Yang, D

# Setup
This week, we're going to talk about regression and t-test (with a little data cleaning in there!)
But first, let's get set up.
```{r}
install.packages("ggeffects")
install.packages("tidyr")
install.packages("Rmisc")

library(ggplot2)
library(ggeffects)
library(tidyr)
library(psych)
library(Rmisc)
library(dplyr)
```

And then we can import some data to look at!
```{r}
df <- read.csv("Week 4 Favorability Data.csv")
```

# Regression
## Computing Regression
This is the same data we were using last time when we found correlations. We can use the same columns (Dem and Rep Favorability) to find a regression:
```{r}
reg <- lm(repFav~demFav, data=df)
summary(reg)
```
This may be a little confusing to read, but know that the "Estimate" is the b-coefficient for your predictor and that the Pr(>|t|) bit is the p-value.
The first "Estimate," by (Intercept), is the intercept!
So we can see that the formula for our regression is Y-hat = 

This is cool, but there's more we can do! We can find confidence intervals around our values:
```{r}
confint(reg)
```

We also are using data that doesn't have a meaningful zero, meaning that we don't get much out of knowing our intercept.
Let's change that:
```{r}
df$cDemFav <- scale(df$demFav,center=T,scale=F)
```

This only centers our data; it doesn't z-score it. But now, if we run another regression:
```{r}
reg2 <- lm(repFav~cDemFav,data=df)
summary(reg2)
```
Now our intercept has meaning!

We can also standardize our data to get beta coefficients in our regression:
```{r}
df$zDemFav <- scale(df$demFav)
df$zRepFav <- scale(df$repFav)

reg3 <- lm(zRepFav~zDemFav,data=df)
summary(reg3)
```

## Plotting Regression
We can then plot our regression, using ggpredict
```{r}
gg <- ggpredict(reg,terms=c("demFav"))

plot(gg)+labs(x="Democrat Favorability",y="Republican Favorability",title="")
```

# t-tests
Now, we're going to talk about t-tests.
## One Sample
First, let's start with a one-sample t-test. We might be wondering if our Republican favorability or Democrat favorability is different from the mean. We can test that:
```{r}
t.test(df$demFav,mu=2.5)
```

## Two Sample
One sample t-tests aren't very common, so we can instead talk about two sample t-tests, using the same code.
We can compare, for example, if people differ in favorability towards Democrats or Republicans. But first, let's check our homogeneity of variance:
```{r}
var.test(df$demFav,df$repFav)
```
Based on this, what do we know? Should we assume the variances are equal?
```{r}
t.test(df$demFav,df$repFav,var.equal=)
```
But now, is an independent samples t-test the right choice?
```{r}
t.test(df$demFav,df$repFav,var.equal=,paired=)
```

There's also another way to do a t-test based on how your data is organized.
Let's say we want to only look at Democrats and Republicans and see how they vary in how they feel about the Republican party. First, we can isolate just them:
```{r}
df1 <- subset(df,df$PARTY=="Republican"|df$PARTY=="Democrat")
```

And then we can run another t-test! Testing variance first:
```{r}
var.test(repFav~PARTY,data=df1)
```

And similarly running a t-test:
```{r}
t.test(repFav~PARTY,data=df1,var.equal=)
```

R output gives you means for the groups, but you may also want standard deviations when writing up your results in APA style.
Here's how you would find standard deviation by group:
```{r}
aggregate(repFav~PARTY, df1, sd)
```

For the t-tests by column (the above ones comparing demFav and repFav), you can just run the standard deviation function on the whole column, as we've done before:
```{r}
sd(df$demFav,na.rm=T)
```
```{r}
sd(df$repFav,na.rm=T)
```


## Graphing t-tests
We may want to graph t-tests as bar graphs with confidence intervals.
```{r}
dfSumm <- df1 %>%
  group_by(PARTY) %>%
  summarise(
    n=n(), #Tells it what n is for SE calculation later
    sd = sd(repFav, na.rm=T), #Finds SD for SE calculation
    repFav = mean(repFav,na.rm=T) #Mean for later
    )%>%
  mutate( se=sd/sqrt(n))  %>% #Calculates SE
  mutate( ci=se * 1.96) #Calculates confidence interval. You may want to add in your t-crit
            
ggplot(dfSumm, aes(PARTY, repFav)) +
  geom_col() +
  geom_errorbar(aes(ymin = repFav-ci, ymax = repFav+ci),width=0.3)+
  xlab("Political Party")+
  ylab("Republican Favorability")
```

# Transforming Data
Sometimes, you may want to store your data in different ways. For t-tests, you can largely use data regardless of if your groups are in separate columns or in the same column with a grouping variable elsewhere, but other measures may require you to change it (or you may just want to).
Two functions that may be especially helpful are pivot_wider and pivot_longer.
```{r}
dfLong <- pivot_longer(df,cols=c("repFav","demFav"),
                    names_to="polParty",values_to="partyFav")
```

I would do this to make a bar chart, because I just taught myself how to make a bar chart with data stored in this way and Google is not helpful in making a bar chart in a different way.
```{r}
dfLong$polParty <- as.factor(dfLong$polParty)

dfSumm2 <- dfLong %>%
  group_by(polParty) %>%
  summarise(
    n=n(), #Tells it what n is for SE calculation later
    sd = sd(partyFav, na.rm=T), #Finds SD for SE calculation
    partyFav = mean(partyFav,na.rm=T) #Mean for later
    )%>%
  mutate( se=sd/sqrt(n))  %>% #Calculates SE
  mutate( ci=se * 1.96) #Calculates confidence interval. You may want to add in your t-crit

ggplot(dfSumm2, aes(polParty, partyFav)) +
  geom_col() +
  geom_errorbar(aes(ymin = partyFav-ci, ymax = partyFav+ci),width=0.3)+
  xlab("Political Party")+
  ylab("Favorability")+
  scale_x_discrete(labels=c("repFav" = "Republican Party", "demFav" = "Democratic Party"))
```

You may also want to do the reverse of this kind of code.
```{r}
dfWide <- pivot_wider(dfLong,names_from="polParty",values_from="partyFav")
```
