---
title: "Week 5"
author: "Emily Esposito"
date: "10/29/2021"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

# Setup
First, let's get set up.
```{r}
install.packages("lsr")

library(lsr)
library(psych)
```

And then we can import some data to look at!
```{r}
df <- read.csv("Week 4 Favorability Data.csv")
```

# T-test Effect Sizes
We can look at the data from last week and calculate effect size (Cohen's d)

Let's first look at the t-test this way, as we did last week.
We can isolate just the Republicans and Democrats:
```{r}
df1 <- subset(df,df$PARTY=="Republican"|df$PARTY=="Democrat")
```

And then we can run another t-test! Testing variance first:
```{r}
var.test(repFav~PARTY,data=df1)
```

And similarly running a t-test:
```{r}
t.test(repFav~PARTY,data=df1,var.equal=F)
```

And then do a measure of effect size:
```{r}
cohen.d(repFav~PARTY,data=df1)
```

This code gives both the Cohen's d and the CI around it.
However, I've had some trouble with this function before, so here's another, just in case:
```{r}
cohensD(repFav~PARTY,data=df1)
```


