---
title: "Week 6"
author: "Emily Esposito"
date: "11/4/2021"
output: html_document
---

# Setup
Load libraries
```{r}
library(lsr)
library(effectsize)
library(tidyr)
library(ggplot2)
library(dplyr)
```

And import data
```{r}
df <- read.csv("Week 4 Favorability Data.csv")
df <- df[!is.na(df$PARTY), ]
```

# Looking at our data
We're going to look at how party affiliation affects attitudes towards Republicans, but first, let's look at the data:
```{r}
df %>%
  ggplot( aes(x=repFav, fill=PARTY)) +
    geom_histogram(position = 'identity') +
    labs(fill="")+
  ylab("Frequency")+
  xlab("Republican Favorability")
```

We may also want to look at mean and standard deviation of our groups
```{r}
aggregate(df$repFav, list(df$PARTY), FUN=mean, na.rm=T)
```

```{r}
aggregate(df$repFav, list(df$PARTY), FUN=sd, na.rm=T)
```


# Running a One-Way ANOVA
Here is the basic formula for running an ANOVA:
```{r}
model1 <- aov(repFav~PARTY,data=df)
summary(model1)
```

And then we can run a few different post hoc tests.

First, we can do pairwise comparisons. Don't forget to do a Bonferroni correction!
```{r}
attach(df)
pairwise.t.test(repFav,PARTY,p.adjust.method = "bonferroni")
```

We can also do Tukey's test
```{r}
TukeyHSD(model1)
```

Or, the much scarier, contrasts.
This code is annoying, because you have to do it semi by hand like we talked about in class.
First, we check our levels:
```{r}
levels(df$PARTY)
```


Then, we write out our contrasts and combine them:
```{r}
c1 <- c(1, -.5, -.5, 0) #Comparing Democrat to Independent+Other
c2 <- c(0, -.5, -.5, 1) #Comparing Republican to Independent+Other

cont <- cbind(c1,c2)

contrasts(df$PARTY) <- cont
contrasts(df$PARTY)
```

And then we run another ANOVA:
```{r}
model2 <- aov(repFav~PARTY,data=df)
summary.aov(model2, split=list(PARTY=list("Dem vs. Ind+Other"=1, "Rep vs. Ind+Other"=2)))
```

Now, we can also do effect size calculations, such as eta-squared:
```{r}
etaSquared(anova,anova=T)
```

Or omega-squared (in the effectsize package):
```{r}
omega_squared(model1)
```

# Graphing ANOVA
A review of bar graphs:
```{r}
ggplot(df,aes(PARTY,repFav))+geom_bar(position="dodge",stat="summary",fun="mean")+xlab("Political Party")+ylab("Mean Republican Favorability")
```